---
title: 'Interacting with APIs: Example using the Financial data API'
author: "Dane Korver"
date: '2022-06-26'
output:
  github_document:
  toc: true
  html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Interacting with APIs: Example using the Financial data API

I am creating a vignette to show how to contact an API using functions I’ve created to query, parse, and return well-structured data.  Then we'll use them to do some exploratory data analysis.

## Requirements 

To use the functions for interacting with the Financial data API, I used the following packages:  

- httr 
- jsonlite 
- tidyverse 
- devtools 

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Load in the packages.
library(httr)
library(jsonlite)
library(tidyverse)
library(devtools)
```

## API Interaction Function

The first challenge with this API was dealing with the variable names.  The default names are not preferable to work with when doing data analysis, so I renamed some of the variables to make them readable.  Once I selected and renamed the variables to make them readable, I wanted to calculate the percentage change between the stock's open and close price.  Now that I have the stock's percentage change, I wanted to output the day's best and worse performers.  I then wanted to look at the stock's opening price at the start and its closing price of the stock at the end of the month and see how many had a experienced a gain, no change or loss for the month.

The `get_stocks` function grabs the daily open, high, low, and close for the entire stocks/equities markets from the API given a date you entered following the Year-Month-Day format, a boolean (true/or false) value for whether you want the prices to be adjusted for splits, and the API key that the service provides you.    

```{r}
get_stocks <- function(day,adjusted,apikey) {
# Set the baseURL, midpoint, endpoint, and combine them with day, adjusted and apikey inputs for the full url.
  baseURL <- "https://api.polygon.io/v2/aggs/grouped/locale/us/market/stocks/"
  midpoint <- "?adjusted="
  endpoint <- "&apiKey="
  fullURL <- paste0(baseURL,day,midpoint,adjusted,endpoint,apikey)
  stocks <- GET(fullURL)
  get_stocks_text <- content(stocks,"text",encoding="UTF-8")
  # Get the API output.
  get_stocks_json <- fromJSON(get_stocks_text, flatten = TRUE)
  # Convert to a dataframe 
  get_stocks_df <- as.data.frame(get_stocks_json)
  # Convert to a tibble
  output <- as_tibble(get_stocks_df)
  output$Date <- day
  # Return the output from the request.
  return(output)
}
```

## Data Exploration

I "pseudo-randomly" chose a month. I write, "pseudo-randomly" because with a free account, you are only able to access certain time points. For this exercise, I chose March, 2021. I grabbed the the March 1, 2021 and March 31, 2021 data from the API.  I selected and renamed the variables that I wanted to use for my analysis and also computed the daily percentage change and outputted a table that lists out the best and worse performers for that day.
```{r echo=FALSE}
get_stocksD030121 <- get_stocks("2021-03-01","FALSE","F9u69d8YyFxs7vSBybGmL8p0hCTGgN_t")
get_stocksD033121 <- get_stocks("2021-03-31","FALSE","F9u69d8YyFxs7vSBybGmL8p0hCTGgN_t")
```
```{r}
stocks_dataD030121 <- get_stocksD030121 %>% 
  select(results.T,results.v,results.o,results.c,results.h,results.l,Date) %>%
  rename(ticker = results.T,volume01 = results.v,open01 = results.o,close01 = results.c,day_high01 = results.h,day_low01 = results.l) %>%
  mutate(daily_pct_change01 = ((close01-open01)/open01)*100) %>%
  arrange(daily_pct_change01)
worst0301<-filter(head(stocks_dataD030121,n=5))
best0301<-filter(tail(stocks_dataD030121,n=5))
best0301<-arrange(best0301,desc(daily_pct_change01))
```
For March 1, here are the top 5 best and worse performers by daily percentage change: 
```{r}
#print out the best and worst performers for March 1, 2021
best0301
worst0301
```

Then for March 31, 2021, I did the same thing:
```{r}
stocks_dataD033121 <- get_stocksD033121 %>% 
  select(results.T,results.v,results.o,results.c,results.h,results.l,Date) %>%
  rename(ticker = results.T,volume31 = results.v,open31 = results.o,close31 = results.c,day_high31 = results.h,day_low31 = results.l) %>%
  mutate(daily_pct_change31 = ((close31-open31)/open31)*100) %>%
  arrange(daily_pct_change31)
worst0331<-filter(head(stocks_dataD033121,n=5))
best0331<-filter(tail(stocks_dataD033121,n=5))
best0331<-arrange(best0331,desc(daily_pct_change31))
```
For March 31, here are the top 5 best and worse performers by daily percentage change: 
```{r}
#print out the best and worst performers for March 31, 2021
best0331
worst0331
```

Then I merged the data together by ticker and computed a monthly percentage change variable and then also created a variable to indicate if the stock had a gain, loss or experienced no change for the month.
```{r}
combine<-inner_join(stocks_dataD030121,stocks_dataD033121,by="ticker")
combine$monthly_pct_change = ((combine$close31-combine$open01)/combine$open01)*100
combine<-arrange(combine,monthly_pct_change)
combine<-select(combine,ticker,volume01,volume31,open01,close31,monthly_pct_change)
combine <- combine %>% mutate(month_gain = if_else(monthly_pct_change>0,"Gain",
                                                   if_else(monthly_pct_change==0,"No     change","Loss")))
month_worst<-filter(head(combine,n=5))
month_best<-filter(tail(combine,n=5))
month_best<-arrange(month_best,desc(monthly_pct_change))
```
Then, for curiosity, I outputted the top 5 best and worse performers for the month.
```{r}
month_best
month_worst
```
Using my `month_gain` variable, I created a contingency table to see how many stocks experienced a gain, loss, or no change during the month of March, 2021.
```{r}
table(combine$month_gain)
```
As we can see by the table, the number of stocks that had a gain or loss by monthly percentage change is almost even and very few had experienced no percentage change for the month.

After investigating the number of stocks that experienced a gain, loss or no change, I wanted to compare the trading volome for these stocks at the start and end of the month.   




